### 2025-11-01 factor-analysis.qmd
20251101 - additions

As a result, to improve the interpretability of the factor analysis, we can do what is called rotation.\index{factor analysis}\index{rotation}
Rotation leverages the idea that there are infinite solutions to the factor analysis model that fit equally well.\index{factor analysis}\index{rotation}
*Rotation* involves changing the orientation of the factors by changing the axes so that variables end up with very high (close to one or negative one) or very low (close to zero) loadings, so that it is clear which factors include which variables.\index{factor analysis}\index{rotation}
That is, rotation rescales the factors and tries to identify the ideal solution (factor) for each variable.\index{factor analysis}\index{rotation}
It searches for simple structure and keeps searching until it finds a minimum.\index{factor analysis}\index{rotation}\index{simple structure}
After rotation, if the rotation was successful for imposing simple structure, each factor will have loadings close to one (or negative one) for some variables and close to zero for other variables.\index{factor analysis}\index{rotation}\index{simple structure}
The goal of factor rotation is to achieve simple structure, to help make it easier to interpret the meaning of the factors.\index{factor analysis}\index{rotation}\index{simple structure}

>>>>

As a result, to improve the interpretability of the factor analysis, we can do what is called rotation.\index{factor analysis}\index{rotation}
Rotation leverages the idea that there are infinite solutions to the factor analysis model that fit equally well.\index{factor analysis}\index{rotation}
*Rotation* involves changing the orientation of the factors by changing the axes so that variables end up with very high (close to one or negative one) or very low (close to zero) loadings, so that it is clear which factors include which variables (and which factors each variable most strongly loads onto).\index{factor analysis!decisions}\index{factor analysis}\index{principal component analysis}\index{rotation}
That is, rotation rescales the factors and tries to identify the ideal solution (factor) for each variable.\index{factor analysis!decisions}\index{factor analysis}\index{principal component analysis}\index{rotation}
Rotation occurs by changing the variables' loadings while keeping the structure of correlations among the variables intact [@Field2012].
Moreover, rotation does not change the variance explained by factors.
Rotation searches for simple structure and keeps searching until it finds a minimum (i.e., the closest as possible to simple structure).\index{factor analysis!decisions}\index{factor analysis}\index{principal component analysis}\index{rotation}\index{simple structure}
After rotation, if the rotation was successful for imposing simple structure, each factor will have loadings close to one (or negative one) for some variables and close to zero for other variables.\index{factor analysis}\index{rotation}\index{simple structure}
The goal of factor rotation is to achieve simple structure, to help make it easier to interpret the meaning of the factors.\index{factor analysis}\index{rotation}\index{simple structure}
